
---
title: "SCC_guide"
description: >
  Somatic Cell Count (SCC) curves fitted to modified Milk Curves with nonlinear models.
output: 
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SCC_guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(woodcurvebox)
```


# Somatic Cell Count (SCC) Guide

Somatic Cell Count (SCC) curves for dairy cows are a useful tool for:

* Surveillance of animal welfare
* Supporting treatment decisions
* Visualize SCC fluctuations throughout a lactation phase to identify potential mastitis cases.

A characteristic of the SCC curve is that it's somewhat similar to a reversed Milk curve.
This means we can fit the SCC observations to modified milk curves, as presented by the Wood's and Wilmink lactation functions.

For fitting, we will use nonlinear models.

We will start by loading the following packages for visualization and modeling:


```{r, message = FALSE}
library(tidyverse)
library(nls.multstart)
library(nlme)
```


# Models
Both Wilmink and Wood's functions can be used to fit SCC. We are using three parameter modified versions


## Wood's function

Nonlog Scale:
SCC = a * DIM ** b * exp(-c * DIM)

Logscale:
log(SCC) = log(a) + b * log(DIM) + (-c * DIM)


## Wilmink function

SCC Logscale:
log(SCC) = a + b * DIM + k * exp(-0.5 * DIM)


# Data Simulation

Data simulating the SCC observations in a 305-day lactation can be simulated using a modified Wilmink's function. Days of milk are set to a length of maximum 305 days, with an interval of 30 days.


```{r}
set.seed(20220329)


simulated_woods_df <- simulate_data(10, 5, 0.1, 2, 0.2, 3, 0.3)

# Estimate parameters
#source("/R/estimate_woods.R") # FIX debug. Cannot source estimate_woods
source("C:/Users/zjt234/PhD/R_Github/woodcurvebox/R/estimate_woods.R") # this workds

# FIX: error below: could not find function "nls.multstart" (library is loaded ?)
estimated_params <- estimate_woods(data = simulated_woods_df) 
# FIX first need to reload nls.multstart
# FIX then need to reload nlmw
# FIX then following error: 
#           Error in f_woods(DIM, loga, b, k) : could not find function "f_woods"

```

## Simulated Data Format

The simulated data frame should follow the variable format required by the woodcurvebox functions:

* Unique animal ID: "cowID"
* Days in Milk (time in lactation): "DIM"
* Somatic Cell Count on a log scale: "logSCC"


# Fitting the Models

## Wood's Style Function

```{r}
# Fitting with nls_multstart
f_woods_nls <- estimate_woods(simulated_woods_df)

# Fitting with nlme
f_woods_nlme <- estimate_woods(simulated_woods_df)

```

## Wilmink Style Function

```{r}
# Fitting with nls_multstart
f_wilmink <- estimate_wilmink(simulated_woods_df)

# Fitting with nlme
f_wilmink <- estimate_wilmink(simulated_woods_df)

```


# Visualizing SCC curves

## SCC Curve - Wood's Style

```{r}
library(ggplot2)

# Plotting Wood's SCC curve
plot_woods <- ggplot(simulated_woods_df, aes(x = DIM, y = logSCC)) +
  geom_point() +
  geom_line(data = simulated_woods_df, aes(x = DIM, y = fitted(f_nls)), color = "blue", linetype = "dashed") +
  geom_line(data = simulated_woods_df, aes(x = DIM, y = fitted(f_nlme)), color = "red", linetype = "dashed") +
  labs(x = "Days in Milk (DIM)", y = "Log SCC") +
  ggtitle("SCC Curve - Wood's Style") +
  theme_minimal()

print(plot_woods)

```

## SCC Curve - Wilmink Style

```{r}
# Plotting Wilmink SCC curve
plot_wilmink <- ggplot(simulated_woods_df, aes(x = DIM, y = logSCC)) +
  geom_point() +
  geom_line(data = simulated_woods_df, aes(x = DIM, y = fitted(f_wilmink)), color = "green", linetype = "dashed") +
  labs(x = "Days in Milk (DIM)", y = "Log SCC") +
  ggtitle("SCC Curve - Wilmink Style") +
  theme_minimal()

print(plot_wilmink)

```


