---
title: "SCC_guide"
description: >
  Somatic Cell Count (SCC) curves fitted to modified Milk Curves with nonlinear models.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SCC_guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(woodcurvebox)
```

Somatic Cell Count (SCC) curves for dairy cows are a useful tool for:

  * Surveillance of animal welfare
  
  * Supporting treatment decisions
  
  * Visualize SCC fluctuations throughout a lactation phase to identify potential masttitis cases.
  
  
A characteristic for the SCC curve is that it's somewhat similar to a reversed Milk curve.
This means we can fit the SCC observations to modified milk curves, this being the Wood's and Wilmink lactation functions.

For fitting, we will use nonlinear models.

We will start by loading following packages for visualization and modelling: 

```{r, message = FALSE}
library(tidyverse)
library(nls.multstart)
library(nlme)
```


# Models
Both Wilmink and Wood's can be used to fit SCC.
Fewer unknown parameters in Wood's, that has 3 compared to the 4 in Wilmink.

FIX: Update with a three parameter wilmink

## Wood's function

Nonlog Scale:
a * DIM ** b * exp(-c * DIM)

Logscale:
log(a) + b * log(DIM) + (- c* DIM)

```{r}
## FIX: Call function from woodcurvebox instead

f_woods <- function(DIM, a,b,c){
  a * DIM ** b * exp(-c * DIM)
}

```

## Wilmink function

Logscale:
a + b * DIM + exp(-(exp(k)) * DIM)*d

 
```{r}
## FIX: Call function from woodcurvebox instead. Both 3 and 4 parameter

f_wilmink <- function(DIM, a,b,k,d){
  a + b * DIM + exp(-(exp(k)) * DIM)*d
}

```

# Data

Data simulating the SCC observations in a 305 days lactation can be simulated using a modified Wood's function.
The 3 unknown parameters have been established through various tests (link to paper II when possible).
Days of milk are set to a length of maximum 305 days, with an interval of 30 days.

```{r}
set.seed(20220329)

woodcurvebox:::simulate_woods

```


First we create the data frame for individual cows

```{r}
# FIX : this should be automated as a function
# FIX: include rnorm variance for each parameter
gen_cow_ids <- str_c("CKR", 1:10)
gen_cow_ids %>%
  map_df(function(cowID) simulate_woods(a = rnorm(1, 5, 0.1), ..., cowID = cowID)) %>%
  identity() -> simulated_woods_df
# simulate_woods(cowID = gen_cow_ids) # doesn't work because `simulate_woods`
# does not vectorize over `cowID`

```


## variable format

Names for variables should be according to the woodcurvebox functions, where:
UNique animal ID : "cowID"
Days in Milk, time in lactation: "DIM"
Somatic Cell Count on a log scale: "logSCC"



# Fitting

## nls multistart

```{r}
f_nls <- 
  nls.multstart::nls_multstart(
  logSCC ~ a * (DIM**b) * exp(-c * DIM),
  data = simulated_woods_df,
  lower=c(a=0, b=0, k=-5, d=0), 
  upper=c(a=9, b=1.5, k=0, d=5),
  start_lower = c(a=0, b=0, k=-5, d=0),
  start_upper = c(a=8, b=1, k=-0.01, d=4),
  iter = 500,
  supp_errors = "Y")

```


## nlme 

```{r}
f_nlme <-
  nlme::nlme(
  #scc ~ a * (DIM**b) * exp(-c * DIM),
  logSCC ~ f_woods(DIM, a, b, c),
  data = data,
  fixed = a + b + c ~ 1,
  random = a + b + c ~ 1,
  # groups = ~cowID,
  # groups = ~herdID,
  start = c(
    a = 150,
    b = -0.1,
    c = 0.003
  ),
  # na.action = na.exclude,
  control = list(returnObject = TRUE)
)
```

# Comparing the models

## Parameters

nls vs nlme.
woods vs wilmink.

GAM ?

Compare the fit with density plot.

Assess the models with ?




## Fit

Prediciton curves for each model



# Additional - fitting on herd level

We include a herd variable, and create noise specific for each herd.

We are adding a herd ID to the data, including noise (epsilon_herd) specific to each herd

```{r}
simulated_woods_df %>%
  dplyr::mutate(herdID = dplyr::if_else(str_detect(cowID, "[1-8]{1,1}$"), "A", "B")) %>%
  dplyr::group_by(herdID) %>%
  # mutate(epsilon_herd = rnorm(n = 1, mean = 0, sd = sd_herd)) %>%
  dplyr::mutate(epsilon_herd = rnorm(n = 1, mean = 0, sd = 2)) %>%
  dplyr::ungroup()

```

## Fitting on herd level

nlme example on herd level
